Theoretical prediction of disrupted Min oscillation in flattened \emph{Escherichia coli}

### 2015-03-22 paper/plos.tex
Continued to address referee issues, recalculated correlation data, although the stochastic shape A is missing, Ill look into that Tested everything on crystal.

% Template for PLoS
% Version 3.0 December 2014
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
% the final text of your manuscript.
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout or alignment
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

@@

\fixme{
  First referee:
  \begin{enumerate}
  \item \blue{Done} Add discussion of ``platform'' --- wrote our own code,
    publically available (cite github?).
  \item \blue{Done} Make 3D geometries available? Cite that the publically
    available code can generate them?
  \item Claims differential equations aren't needed.
  \item Claims wild-type movie-ish plot not needed.
  \item Points out problems with Huang 2003:
    \begin{enumerate}
    \item \blue{David.  I've added some sentences discussing the
      starry night in the context of the membrane diffusion.  I didn't
      say anything about a one half period allowing diffusion to smear
      an entire pole, because it seems to contradict the second part
      of this paragraph.  What do you think?} Lacks diffusion on the
      membrane. Supported by Meacci 2006, which experimentally tries
      to measure diffusion on membrane, has a tough time, but shows
      that definitely diffuses.
    \item Underestimates diffusion in cytoplasm by a factor of 5-6.5.
      Supported by Meacci 2006, who concludes a rate of 16 microns
      squared per sec for MinD and 10 microns squared per sec for
      MinE.
    \item Uses MinD-ADP -> MinD-ATP that is 3 times lower than
      experimental bound. Halatek uses ADP->ATP of 1/6 s-1. They cite
      Meacci 2006, who state that there is a lower bound on reference
      time of 3 seconds, so this constant has an upper bound of 1/3
      s-1.
    \item (Still debated) No evidence that DE recruits to membrane.
    \item Works in only a limited part of parameters phase space, and
      could not account for important things like
      temperature-independence.
    \end{enumerate}
  \item \blue{Done} Height should be bigger to compare with experiment.  0.4-0.6
    microns.  The 0.25 micron height of the cells used in this study
    does not represent the experimental settings well. The simulations
    are based on the flattened cell shapes reported from the
    experiments by Mannik et al 2009 and 2012. In these experimental
    studies, E. coli were shown to be able to grow in channels as low
    as 0.6 micron in hard silicon structures, and no more than 0.4
    micron (see Mannik 2009). Indeed, 0.25 micron channels were
    constructed with PDMS ceilings, which is a rather flexible polymer
    allowing bacteria to deform it to effectively higher
    channels. Thus the simulations would be more realistic if
    performed in the range of 0.4-0.6 micron thickness.

    \emph{If you can't make it grow in a .3 micron (rigid) silicon channel,
      then we and you observe it growing in a .25 micron flexible
      channel, it presumably must be able to grow because it deformed
      the channel to larger than .3 microns.}
  \item \blue{David} Starry night in the wild-type cells is a problem.
  \item Better explanation of correlation function (which is
    interesting!).
  \item Shape A is the interesting bit in our main figure.
  \item Would like more analysis of stochastic locations in different
    geometries.  Clusters?  Could add more geometries?
  \item Bold conclusions about flattening.  Talk about both width and
    thickness.  Compare Halatek 2012.  Wants systematically more
    geometries again.
  \item Is doubtful of our correlation plots relative to those of
    Fange and Elf 2006.
  \item Clarify (esp in conclusions) what we mean by flattening and
    irregular shape.  ``breaking the cylindrical symmetry and
    flattening the cells.'' versus ``breaking the mirror plane
    symmetry of the cross-section of a flattned cell.''
  \item[TODO suggestion] Switch to Halatek and Frey
  \item[TODO suggestion] Thoroughly explor shape parameter space
  \item[TODO suggestion] Check robustness of our conclusions with
    respect to changes in (reaction-diffusion?) parameters.
  \item Check order of figures and their references.
  \item Cite biological (non-computational) works e.g. Corbin 2008.
  \end{enumerate}
  Second referee:
  \begin{enumerate}
  \item Nitpicks about language of deterministic model versus
    stochastic model.  (Read carefully)
  \item \blue{David} Explain in greater detail the obviousness (maybe) that the
    deterministic approach doesn't have fluctuations.  RESPOND:
    chaotic systems do deterministically result in fluctuations.
    Explain this somewhere, and discuss triviality of the conclusion
    that deterministic is wrong because it does not exhibit these
    irregularities.
  \item Would like temporal averages over cell-cycle time scales.
  \item More discussion of how flattening leads to spatial
    irregularities.  Compare with large nearly spherical e. coli Shih
    2005.  See also Halatek and Frey 2012 for rotating oscillation
    axis due to destabilization of short modes.  How does this connect
    with our stability analysis?
  \item \blue{David} Should mention more recent revisions of the model.
  \item Fange and Elf do not use the same reaction-diffusion
    equations.
  \item Clarify statement about 1000 proteins and stochastic
    fluctuation.  Reword about a continuum description omitting
    information?
  \item \blue{Done} Do not call Huang parameters ``wild type values.''  They are
    fitted parameters.
  \item Add methods section for numerics (deterministic).
  \item \blue{Done} Add details of the linear stability analysis.
  \item \blue{David} Still need to fix periods on correlation plots.
    \blue{About this point - I updated the data and it seems the
      result is defferent from before - the stadium shapes seem to be
      temporally correlated for much longer compared with the shape A
      than before.}
  \end{enumerate}
}

@@

However, Mannik \emph{et al.} have recently shown that there are
limitations to this robust capability to
oscillate~\cite{mannik2012robustness,mannik2009bacterial,mannik2010bacteria}. They
have experimentally forced \emph{E. coli} cells into microfabricated
channels of thickness less than that of the natural diameter of the
cells, in which they area able to grow and divide. Upon entering the
channels, the cells undergo a mechanical deformation in which they
both widen and lengthen within the plane of the channel.  This
deformation results in very wide (they can reach widths of over
$5\micron$) flattened cells that when viewed from the top down have
irregular and asymmetric shapes.  While these cells are still able to
divide into surprisingly equal volumes, the MinD oscillations in these
cells are irregular, both temporally and spatially. Seen with
fluorescent microscopy, the MinD maximize in multiple locations within
the cell, in a seemingly random sequence. These experiments allow for
an opportunity to test MinD simulation models against more extreme
cases than have been seen thus far.

A number of models of the MinD protein system have been developed that
accurately describe its basic oscillatory nature.
%
Early models involved free proteins that affect each others' rates of
diffusion and membrane attachment, but do not combine into compound
states~\cite{meinhardt2001pattern}.  In 2003 Huang improved upon this
approach with a simple and very successful simulation model based on
MinD-MinE combination, ATPase hydrolysis, and MinD membrane attachment
that exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is
recruited to the membrane by MinD that is already clustered there
(following observed non-linear attachment of MinD on the cell
membrane~\cite{hu2002dynamic,shih2002division}), and is stationary
once attached.
%
\blue{(5)} Several models
\cite{fange2006noise,bonny2013membrane,halatek2012highly} modify
Huang's original model to more accurately treat experimental findings
published after his 2003 paper
\cite{meacci2006mobility,loose2011min}. These models introduce slow
diffusion along the membrane (two to three orders of magnitude slower
than that within the cytoplasm), revise the process of MinD
recruitment to the membrane, and change some of the constants used in
the model. The Bonny 2013 model assumes an instantaneous nucleotide
exchange rate, and allows MinE to remain independently attached to the
membrane after dissasociating with MinD~\cite{bonny2013membrane}.
%

%%  A number of studies have used an approach similar in
%% that they do not rely on the ability of MinD to move along the walls
%% and cluster~\cite{kruse2007experimentalist, meinhardt2001pattern,
%%   drew2005polymerization, fange2006noise, kerr2006division}, while
%% studies have been made as well of models which rely on MinD mobility
%% and attraction on the cell membrane~\cite{kruse2002dynamic,
%%   howard2005cellular}.

There are only on the order of 1000 MinD proteins in a given cell,
which means that stochastic fluctuation can be expected to play a
large role.  To treat this, there have been a number of variations of
the Huang 2003 model that stochastically simulate the same \fixme{similar}
reaction-diffusion equations~\cite{fange2006noise, kerr2006division}.
These studies largely confirm the results of Huang's deterministic
model when applied to the wild-type, pill-shaped phenotype.  However,
the stochastic models are slightly more successful in predicting
experimentally observed oscillations in round cell
phenotypes~\cite{fange2006noise, huang2004min}, and they enable
prediction of fluctuations in the predicted
behavior~\cite{kruse2007experimentalist}.  Both deterministic and
stochastic models are widely used throughout systems
biology~\cite{lawson2013spatial, robb2014stochastic,
  oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
have unique advantages and limitations.  The deterministic approach
has an advantage in providing a simple prediction of average behavior,
while the stochastic approach enables prediction of fluctuations from
that mean, and reduces sensitivity to initial conditions.

%% biology~\cite{lawson2013spatial, robb2014stochastic,
%%   oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
In this paper we use Huang's original 2003
model~\cite{huang2003dynamic}, which has been widely
used~\cite{hattne2005stochastic} \fixme{search for more} in both
deterministic and stochastic variants, to study flattened cells of
$0.4\micron$ thickness that are similar to those observed by Mannik
\emph{et al.}~\cite{mannik2009bacterial}.  \fixme{Some weak
  justification... it has been widely used.}

@@

Huang's simulations use total MinD and MinE concentrations of
$1,000/\micron$ and $350/\micron$, respectively, in a cylindrical cell
of radius $0.5\micron$, and in our (non-cylindrical) cells we use the
same number of proteins per unit volume.  These concentration values
are $1273\micron^{-3}$ and $446\micron^{-3}$,
respectively. \blue{(1,2)} We have chosen to write our own simulation
platform and membrane creation tools from scratch, instead of using
existing available software used in previous studies\fixme{cite
  previous studies} (we believe there is value in re-creating results
with alternative software).  Our simulations take place within a
three-dimensional Cartesian grid that has a grid spacing of
.05\micron, and our cell shapes are specified as the zero of an
analytic function which can generate the geometry presented in this
paper.  Our code is publically available on github.com~\fixme{cite
  github account in references? Yes!}

@@

\blue{(5a),(7)} Both Fig.~\ref{image-p} and its associated movies (see
supplementary material), show that the stochastic model exhibits a
``starry night'' effect characterized by spatially fixed points of
protein density build up, as recruitment leads to clusters of MinD
forming on the wall and then subsequently dissipating.  We attribute
this effect to the fact that the Huang model does not include
diffusion along the membrane, which has been experimentally verified
yet found to be two orders of magnitude less than diffusion in the
cytoplasm~\cite{meacci2005min}. Even ignoring this effect, however, it
is clear from Fig.~\ref{image-p} that the stochastic simulation
results in a protein distribution that is less spatially regular than
the deterministic model predicts.  This naturally leads to
irregularity in the temporal periodicity of the system, which we
quantify by examining the temporal correlation function of the total
MinD found in two opposite polar regions.
%
This correlation function, which is displayed in Fig.~\ref{corr-pill},
is given by
\begin{equation}
  C(\tau) \propto \int
  (N_{\textit{top}}(t) - \bar N_{\textit{top}})
  (N_{\textit{bottom}}(t+\tau) - \bar N_{\textit{bottom}})dt
\end{equation}
where $N_{\textit{top}}(t)$ and $N_{\textit{bottom}}(t)$ are the total
MinD proteins in the top and bottom thirds of a cell.  These plots
help us in studying the periods and the regularity and stability of
oscillations.  These curves are well fit with a simple decoherence
model, given by the equation
\begin{equation}
  C(\tau) = -\cos\left(\frac{2\pi\tau}{T}\right) e^{-\frac{\tau}{\tau_c}}
\end{equation}
where $T$ is the period and $\tau_c$ is the coherence time.  In the
case of the stochastic model the pill exhibits coherence times of
about 8 periods and in the case of the deterministic model oscillate
indefinitely with complete coherence, indicating that its behavior is
perfectly periodic.

@@

In every case, including the wild-type pill shape, we see irregularity
in the location of the maxima when using the stochastic model.  The
deterministic model shows uniformly bipolar oscillation. \blue{(2)}
It's worth pointing out that this is a significant result, since for
irregular shapes it would be reasonable to hypothesize a chaotic
system, which would be shown by irregular determinisitic data.  From
these results and how they compare with experiment, we conclude that
the deterministic model is inadequate to explain experimental
observations of the locations of density maxima of the MinD protein.

@@

\section*{Conclusion}
We find agreement between our stochastic simulations and experiments
showing disrupted bipolar MinD oscillation in asymmetric flattened
E. coli cells~\cite{mannik2012robustness}.  As observed
experimentally, our simulations predict that MinD maxima will form in
a spatially irregular sequence of locations.  This result builds upon
existing results showing that this model is effective in a variety of
wild-type and mutated cell shapes~\cite{fange2006noise, varma2008min,
  kruse2007experimentalist}, and reinforces that the stochastic
variation of Huang's 2003 model~\cite{fange2006noise,
  kerr2006division} has considerable predictive power beyond the
standard wild-type cell.  \fixme{Discuss which of these are likely to
  map to more recent models.}

>>>>

% Template for PLoS
% Version 3.0 December 2014
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
% the final text of your manuscript.
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout or alignment
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment. For example , CO2 will be CO\textsubscript{2}.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

@@

\fixme{
  First referee:
  \begin{enumerate}
  \item \blue{Done} Add discussion of ``platform'' --- wrote our own code,
    publically available (cite github?).
  \item \blue{Done} Make 3D geometries available? Cite that the publically
    available code can generate them?
  \item Claims differential equations aren't needed.
  \item Claims wild-type movie-ish plot not needed.
  \item Points out problems with Huang 2003:
    \begin{enumerate}
    \item \blue{David.  I've added some sentences discussing the
      starry night effect and membrane diffusion blurring out the
      stars.  I have some confusion however about the diffusion
      lengths.}  Lacks diffusion on the membrane. Supported by Meacci
      2006, which experimentally tries to measure diffusion on
      membrane, has a tough time, but shows that definitely diffuses.
    \item \blue{David} Underestimates diffusion in cytoplasm by a factor of 5-6.5.
      Supported by Meacci 2006, who concludes a rate of 16 microns
      squared per sec for MinD and 10 microns squared per sec for
      MinE.
    \item \blue{David} Uses MinD-ADP -> MinD-ATP that is 3 times lower than
      experimental bound. Halatek uses ADP->ATP of 1/6 s-1. They cite
      Meacci 2006, who state that there is a lower bound on reference
      time of 3 seconds, so this constant has an upper bound of 1/3
      s-1.
    \item \blue{David} (Still debated) No evidence that DE recruits to membrane.
    \item Works in only a limited part of parameters phase space, and
      could not account for important things like
      temperature-independence.
    \end{enumerate}
  \item \blue{Done} Height should be bigger to compare with experiment.  0.4-0.6
    microns.  The 0.25 micron height of the cells used in this study
    does not represent the experimental settings well. The simulations
    are based on the flattened cell shapes reported from the
    experiments by Mannik et al 2009 and 2012. In these experimental
    studies, E. coli were shown to be able to grow in channels as low
    as 0.6 micron in hard silicon structures, and no more than 0.4
    micron (see Mannik 2009). Indeed, 0.25 micron channels were
    constructed with PDMS ceilings, which is a rather flexible polymer
    allowing bacteria to deform it to effectively higher
    channels. Thus the simulations would be more realistic if
    performed in the range of 0.4-0.6 micron thickness.

    \emph{If you can't make it grow in a .3 micron (rigid) silicon channel,
      then we and you observe it growing in a .25 micron flexible
      channel, it presumably must be able to grow because it deformed
      the channel to larger than .3 microns.}
  \item \blue{David} Starry night in the wild-type cells is a problem.
  \item Better explanation of correlation function (which is
    interesting!).
  \item Shape A is the interesting bit in our main figure.
  \item Would like more analysis of stochastic locations in different
    geometries.  Clusters?  Could add more geometries?
  \item Bold conclusions about flattening.  Talk about both width and
    thickness.  Compare Halatek 2012.  Wants systematically more
    geometries again.
  \item Is doubtful of our correlation plots relative to those of
    Fange and Elf 2006.
  \item \blue{David. Read the refferee report with regard to this, and
    not sure what he wants...} Clarify (esp in conclusions) what we
    mean by flattening and irregular shape.  ``breaking the
    cylindrical symmetry and flattening the cells.'' versus ``breaking
    the mirror plane symmetry of the cross-section of a flattned
    cell.''
  \item[TODO suggestion] Switch to Halatek and Frey
  \item[TODO suggestion] Thoroughly explor shape parameter space
  \item[TODO suggestion] Check robustness of our conclusions with
    respect to changes in (reaction-diffusion?) parameters.
  \item Check order of figures and their references.
  \item \blue{David.  I cited the Corbin paper (although he's got a
    typo because Corbin 2008 is after the fange 2006 paper. I could
    cite other experimental papers, but the sentence in question
    specifically refers to changes made to the Huang model, and the
    papers cited are the main reasons that those changes have been
    made.)} Cite biological (non-computational) works, that the
    e.g. Corbin 2008.
  \end{enumerate}
  Second referee:
  \begin{enumerate}
  \item Nitpicks about language of deterministic model versus
    stochastic model.  (Read carefully)
  \item \blue{David} Explain in greater detail the obviousness (maybe) that the
    deterministic approach doesn't have fluctuations.  RESPOND:
    chaotic systems do deterministically result in fluctuations.
    Explain this somewhere, and discuss triviality of the conclusion
    that deterministic is wrong because it does not exhibit these
    irregularities.
  \item Would like temporal averages over cell-cycle time scales.
  \item More discussion of how flattening leads to spatial
    irregularities.  Compare with large nearly spherical e. coli Shih
    2005.  See also Halatek and Frey 2012 for rotating oscillation
    axis due to destabilization of short modes.  How does this connect
    with our stability analysis?
  \item \blue{David} Should mention more recent revisions of the model.
  \item \blue{David} Fange and Huang do not use the same reaction-diffusion
    equations.
  \item \blue{David.  I read the referee's note about this.  Not sure
    what he/she wants.}  Clarify statement about 1000 proteins and
    stochastic fluctuation.  Reword about a continuum description
    omitting information?
  \item \blue{Done} Do not call Huang parameters ``wild type values.''  They are
    fitted parameters.
  \item \blue{Done?  I think our additions to the Methods section
    addressed this?}  Add methods section for numerics
    (deterministic).
  \item \blue{Done} Add details of the linear stability analysis.
  \item \blue{David} Still need to fix periods on correlation plots.
    \blue{About this point - I updated the data and it seems the
      result is defferent from before - the stadium shapes seem to be
      temporally correlated for much longer compared with the shape A
      than before.}
  \end{enumerate}
}

@@

However, Mannik \emph{et al.} have recently shown that there are
limitations to this robust capability to
oscillate~\cite{mannik2012robustness,mannik2009bacterial,mannik2010bacteria}. They
have experimentally forced \emph{E. coli} cells into microfabricated
channels of thickness less than that of the natural diameter of the
cells, in which they area able to grow and divide. Upon entering the
channels, the cells undergo a mechanical deformation in which they
both widen and lengthen within the plane of the channel.  This
deformation results in very wide (they can reach widths of over
$5\micron$) flattened cells that when viewed from the top down have
irregular and asymmetric shapes.  While these cells are still able to
divide into surprisingly equal volumes, the MinD oscillations in these
cells are irregular, both temporally and spatially. Seen with
fluorescent microscopy, the MinD maximize in multiple locations within
the cell, in a seemingly random sequence. These experiments allow for
an opportunity to test MinD simulation models against more extreme
cases than have been seen thus far.  \blue{(2)} An additional
interesting question is whether the irregularity observed by Mannik
\emph{et al.} indicates that the MinD system can become chaotic (as in
e.g. Ref.~\citenum{kuramoto1978diffusion}), or if this irregularity is
the result of a stochastic but nonchaotic process.
%
A number of models of the MinD protein system have been developed that
accurately describe its basic oscillatory nature.
%
Early models involved free proteins that affect each others' rates of
diffusion and membrane attachment, but do not combine into compound
states~\cite{meinhardt2001pattern}.  In 2003 Huang improved upon this
approach with a simple and very successful simulation model based on
MinD-MinE combination, ATPase hydrolysis, and MinD membrane attachment
that exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is
recruited to the membrane by MinD that is already clustered there
(following observed non-linear attachment of MinD on the cell
membrane~\cite{hu2002dynamic,shih2002division}), and is stationary
once attached.
%
\blue{(5a,b,c,d)} Several models
\cite{fange2006noise,bonny2013membrane,halatek2012highly} modify
Huang's original model to more accurately treat experimental findings,
some of which were published after his 2003 paper\blue{(15)}
\cite{meacci2006mobility,loose2011min,corbin2002exploring}. These
models introduce slow diffusion along the membrane (two to three
orders of magnitude slower than that within the cytoplasm), revise the
process of MinD recruitment to the membrane, increase the rate of
diffusion within the cytpolasm, and decrease or assume
instanaeous~\cite{bonny2013membrane} the nucleotide exchange rate.  In
addition, Bonny \emph{et al.} allows MinE to remain independently
attached to the membrane after dissasociating with
MinD~\cite{bonny2013membrane}.
%
There are only on the order of 1000 MinD proteins in a given cell,
which means that stochastic fluctuation can be expected to play a
large role.  To treat this, there have been a number of variations of
the Huang 2003 model that stochastically simulate the similar
reaction-diffusion equations~\cite{fange2006noise, kerr2006division}.
These studies largely confirm the results of Huang's deterministic
model when applied to the wild-type, pill-shaped phenotype.  However,
the stochastic models are slightly more successful in predicting
experimentally observed oscillations in round cell
phenotypes~\cite{fange2006noise, huang2004min}, and they enable
prediction of fluctuations in the predicted
behavior~\cite{kruse2007experimentalist}.  Both deterministic and
stochastic models are widely used throughout systems
biology~\cite{lawson2013spatial, robb2014stochastic,
  oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
have unique advantages and limitations.  The deterministic approach
has an advantage in providing a simple prediction of average behavior,
while the stochastic approach enables prediction of fluctuations from
that mean, and reduces sensitivity to initial conditions.

%% biology~\cite{lawson2013spatial, robb2014stochastic,
%%   oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
In this paper we use Huang's original 2003
model~\cite{huang2003dynamic}, which has been widely
used~\cite{hattne2005stochastic,huang2004min,kerr2006division}
\fixme{I couldn't find too many.  Most (like Bonny and Halatek) use
  modified versions.} in both deterministic and stochastic variants,
to study flattened cells of $0.4\micron$ thickness that are similar to
those observed by Mannik \emph{et al.}~\cite{mannik2009bacterial}.
\fixme{Some weak justification... it has been widely used.}

@@

Huang's simulations use total MinD and MinE concentrations of
$1,000/\micron$ and $350/\micron$, respectively, in a cylindrical cell
of radius $0.5\micron$, and in our (non-cylindrical) cells we use the
same number of proteins per unit volume.  These concentration values
are $1273\micron^{-3}$ and $446\micron^{-3}$,
respectively. \blue{(1,2)} We have chosen to write our own simulation
platform and membrane creation tools from scratch, instead of using
existing available software used in previous
studies~\cite{bonny2013membrane,fange2006noise,halatek2012highly,kerr2006division}
(we believe there is value in re-creating results with alternative
software).  Our simulations take place within a three-dimensional
Cartesian grid that has a grid spacing of .05\micron, and our cell
shapes are specified as the zero of an analytic function which can
generate the geometry presented in this paper.  Our code is publically
available on github.com~\cite{git}.

@@

\blue{(5a),(7)} Both Fig.~\ref{image-p} and its associated movies (see
supplementary material), show that the stochastic model exhibits a
``starry night'' effect characterized by spatially fixed points of
protein density build up, as recruitment leads to clusters of MinD
forming on the wall and then subsequently dissipating.  This results
from the lack of inclusion of diffusion along the membrane in Huang's
model, which would spread out these recruitment clusters.  Experiments
have confirmed that diffusion along the membrane does occur, albeit
with a rate two orders of magnitude lower than that in the
cytoplasm~\cite{meacci2005min}.  The 'stars' in the effect seem to
last for about 10 seconds. The membrane-bound proteins will travel an
approximate distance of \fixme{see note} microns during one of these
periods of time, enough to 'blur out' this effect in wild-type pill
cells.  \fixme{calculating with $<x^2> = Dt$ and $D_{membrane} =
  .16\mu^2/s m$, which is just the experimental cytoplasmic value
  divided by 100, yields 1.2 microns, which maybe is too high for what
  we're trying to say?  Halatek uses $D_{membrane} = .013\mu^2/s m$,
  for which the same calculation would give .36 microns, which is
  better.  Halatek doesn't say why he uses a value that seems to be
  different than what Meacci claim to have found.  Halatek just says
  it's experimentally valid and cites their work.  What to do?}

We also analyze the temporal periodicity of the system, which we
quantify with temporal correlation function of the total MinD found in
two opposite polar regions.  This correlation function, which is
displayed in Fig.~\ref{corr-pill}, is given by
\begin{equation}
  C(\tau) \propto \int
  (N_{\textit{top}}(t) - \bar N_{\textit{top}})
  (N_{\textit{bottom}}(t+\tau) - \bar N_{\textit{bottom}})dt
\end{equation}
where $N_{\textit{top}}(t)$ and $N_{\textit{bottom}}(t)$ are the total
MinD proteins in the top and bottom thirds of a cell.  These plots
help us in studying the periods and the regularity and stability of
oscillations.  These curves are well fit with a simple decoherence
model, given by the equation
\begin{equation}
  C(\tau) = -\cos\left(\frac{2\pi\tau}{T}\right) e^{-\frac{\tau}{\tau_c}}
\end{equation}
where $T$ is the period and $\tau_c$ is the coherence time.  In the
case of the stochastic model the pill exhibits coherence times of
about 8 periods and in the case of the deterministic model oscillate
indefinitely with complete coherence, indicating that its behavior is
perfectly periodic.

@@

In every case, including the wild-type pill shape, we see irregularity
in the location of the maxima when using the stochastic model.  The
deterministic model shows uniformly bipolar oscillation.  \blue{(2)}
We conclude that the model is \emph{not} chaotic (which would show
irregularity in the deterministic computation), but rather that the
irregularity results purely from stochastic processes.  From these
results and how they compare with experiment, we conclude that the
deterministic model is inadequate to explain experimental observations
of the locations of density maxima of the MinD protein.

@@

\section*{Conclusion}
We find agreement between our stochastic simulations and experiments
showing disrupted bipolar MinD oscillation in asymmetric flattened
E. coli cells~\cite{mannik2012robustness}.  As observed
experimentally, our simulations predict that MinD maxima will form in
a spatially irregular sequence of locations.  \blue{(6)}\fixme{Do you think this
  sentence make it sounds like Fange has the same model as Huang?}
This result builds upon existing results showing that this model and
its derivatives are effective in a variety of wild-type and mutated
cell shapes~\cite{fange2006noise, varma2008min,
  kruse2007experimentalist}, and reinforces that stochastic variations
of Huang's 2003 model~\cite{fange2006noise, kerr2006division} has
considerable predictive power beyond the standard wild-type cell.
\fixme{Discuss which of these are likely to map to more recent
  models.}
